% CouncilHound Development Impact Methodology
% Build: tectonic docs/impact-methodology-report.tex
\documentclass[11pt]{article}

\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{array}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\setlist{itemsep=2pt, topsep=4pt}

\newcommand{\assum}[1]{\texttt{#1}}
\title{CouncilHound Development Impact Methodology}
\author{Hunter Brown \\ \small\href{mailto:huntermbrown@gmail.com}{\texttt{huntermbrown@gmail.com}} \\ \small CouncilHound --- impact analysis subsystem}
\date{July 2026 \\ \small model version: joint destination--mode Huff, census-block weights}

\begin{document}
\maketitle

\begin{abstract}
\noindent
CouncilHound produces screening-level impact estimates for proposed local development
projects. A deterministic economic module (spatial-interaction retail capture,
walking activity, an employment ledger) and a deterministic fiscal module
(revenue and service-cost ranges) generate every number; a language model is used
only to extract project facts from public documents under a verbatim-quote
firewall and to write a narrative that a validator constrains to the structured
outputs. This note documents each displayed metric's formula, its academic or
statistical basis, the assumptions that drive its uncertainty range, and the
implementation decisions behind the current system. The estimates rank likely
orders of magnitude under stated assumptions; they are decision-support
evidence, not forecasts.
\end{abstract}

\section{Scope and design principles}

\begin{enumerate}
\item \textbf{Deterministic quantities.} All numbers are produced by closed-form
      models over open data. The language model never supplies a quantity: extracted
      project facts require a verbatim source quote ($\leq$15 words) that is verified
      against the document text in code, and fields without one are nulled. The
      narrative report is post-validated so that every number it contains matches a
      published metric, a stated assumption, a provenance string, or a deterministic
      module note.
\item \textbf{Assumption transparency.} Every contestable parameter is a named
      assumption with a central value, low/high bounds, and a recorded basis. Bounds
      propagate through all formulas by interval arithmetic, so each metric carries a
      range derived from the same assumptions it cites.
\item \textbf{Fail loudly.} Missing inputs (an unpinned tax rate, an absent
      assessment record) degrade the affected metric to ``not computed --- reason,''
      never to a silent default.
\item \textbf{Screening honesty.} Capture, walking, and foot-traffic figures are
      screening estimates in the spatial-interaction tradition; the report states
      this in plain language and names the most sensitivity-driving assumptions.
\end{enumerate}

\section{Notation and data inputs}

\begin{tabular}{ll}
$U$        & proposed dwelling units (extracted from project documents) \\
$\rho$     & occupancy rate at stabilization (\assum{occupancy\_rate}) \\
$\bar{s}$  & persons per occupied multifamily unit (\assum{avg\_hh\_size\_multifamily}) \\
$y$        & mean household income of the site's census tract (ACS B19025/B11001) \\
$\pi$      & new-construction income premium (\assum{income\_premium\_new\_construction}) \\
$e_c$      & CES average annual expenditure for category $c$ (BLS CES 2023) \\
$y_{\mathrm{CES}}$ & CES average pre-tax income per consumer unit (\$101{,}805, 2023) \\
$\varepsilon_c$ & expenditure--income elasticity for category $c$ (Engel scaling) \\
$\kappa$   & CES scaling factor (\assum{ces\_scale}) \\
$A_j$      & attractiveness of destination $j$ (POI count~$=1$ per business) \\
$t^w_j,\ t^d_j$ & walk / drive travel time (minutes) from the site to $j$ \\
$\beta_w,\ \beta_d$ & walk / drive impedance decay per minute \\
$w_c$      & zero-impedance walk preference for category $c$ \\
$\tau$     & daily walk trips per resident (\assum{walk\_trips\_per\_resident\_day}) \\
\end{tabular}

\medskip
\noindent
Abbreviations used throughout: ACS is the Census Bureau's American Community
Survey; CES is the U.S. Bureau of Labor Statistics (BLS) Consumer Expenditure
Survey \cite{blsces}; NHTS is the Federal Highway Administration's (FHWA)
National Household Travel Survey \cite{nhts}; POI is point of interest.

\medskip
\noindent
Networks are OpenStreetMap (OSM) walk and drive graphs built with OSMnx
\cite{boeing2017}; walk edge times use 4.8\,km/h. Population weights use 2020
Census Public Law (PL) 94-171 block centroids. Points of interest merge
Overture Maps, OpenStreetMap, and (when licensed) Foursquare Open Source
Places into a nine-class taxonomy. Travel times come from single-source Dijkstra runs over edge minutes;
each business inherits the time of its nearest network node.

\section{Demand module}

\subsection{New households and residents}
\begin{align}
N_{hh}  &= U \cdot \rho \\
N_{res} &= N_{hh} \cdot \bar{s}
\end{align}
Occupancy at stabilization is benchmarked against the Census Housing Vacancies
and Homeownership (HVS) series \cite{hvs}. Household size uses a multifamily
(MF) bedroom-mix demographic multiplier in the tradition of Listokin et
al.~\cite{listokin2006} ($\bar{s} = 1.8$, range 1.5--2.2, for a typical
studio/one-bedroom/two-bedroom mix): the ACS B25010 tenure-level renter average
($\approx 3.1$ locally) was evaluated and rejected as an estimator because it
is dominated by single-family renter households and is blind to unit mix.
When a project's bedroom mix is known, the multiplier should be set from it
(\S\ref{sec:limitations}).

\subsection{Aggregate income and category spending}
\begin{align}
Y   &= N_{hh} \cdot y \cdot \pi, \qquad r = \frac{y\,\pi}{y_{\mathrm{CES}}} \\
S_c &= N_{hh} \cdot e_c \cdot r^{\,\varepsilon_c} \cdot \kappa
\label{eq:engel}
\end{align}
The tract income $y$ is the \emph{mean} (ACS aggregate household income
B19025 divided by households B11001, summed over the site tract's block
groups) --- the correct estimator for total purchasing power, since
median~$\times$~count understates aggregate income under right-skewed
distributions. When the aggregate tables are suppressed the model falls back
to the population-weighted median and records the substitution. Category
spending uses BLS Consumer Expenditure Survey 2023 levels \cite{blsces}
scaled by the income ratio $r$ raised to a per-category expenditure
elasticity $\varepsilon_c$ (grocery 0.45, restaurant/bar 0.85, comparison
0.95, convenience 0.60, personal services 0.90, entertainment 1.05),
consistent with the Engel gradients in the CES income-quintile tables:
necessity spending grows sublinearly with income, so a high-income tract no
longer implies proportionally more grocery spending.

\section{Retail capture: joint destination--mode choice}
\label{sec:capture}

Retail capture follows the probabilistic trade-area tradition of Huff
\cite{huff1963, huff1964} with the exponential impedance of the
entropy-maximizing spatial-interaction family \cite{wilson1971}. Destination
and mode are chosen \emph{jointly}, as a multinomial logit \cite{mcfadden1974,
benakiva1985} over (business, mode) pairs. For category $c$, business $j$, and
mode $m \in \{w, d\}$:
\begin{equation}
P_c(j, m) \;=\;
\frac{A_j^{\alpha}\, \omega_{c,m}\, e^{-\beta_m t^m_j}}
     {\sum_k A_k^{\alpha} \left( \omega_{c,w}\, e^{-\beta_w t^w_k}
       + \omega_{c,d}\, e^{-\beta_d t^d_k} \right)},
\qquad
\omega_{c,w} = w_c,\ \ \omega_{c,d} = 1 - w_c,
\label{eq:joint}
\end{equation}
with $\alpha = 1$. The walk impedance $\beta_w = 0.10$ min$^{-1}$ is anchored
to empirical negative-exponential decay fits: purpose-specific walking
estimates of 0.09--0.11 min$^{-1}$ (shopping 0.094, restaurant 0.093) from
Twin Cities travel-survey data \cite{iacono2008}, and 0.073 min$^{-1}$ from
the national 2009 NHTS \cite{yang2012}; the sensitivity band (0.05--0.15)
spans the national-sample floor to a hostile-pedestrian-environment ceiling.
The mode preference $w_c$ has a clean interpretation: at
zero travel time for both modes, the walk share is exactly $w_c$; as walk time
grows, walking loses utility to driving and the trip (and its dollars) arrives
by car instead. Annual capture and walk-arriving capture at business $j$:
\begin{equation}
C_j = \sum_c S_c \sum_m P_c(j, m), \qquad
C^{walk}_j = \sum_c S_c \, P_c(j, w).
\end{equation}
Two invariants are pinned by unit tests: $\sum_j C_j = \sum_c S_c$ (joint
probabilities sum to one), and $\sum_j C^{walk}_j \leq \sum_c S_c w_c$ with
equality only when every destination is at zero travel time --- the walk total
is \emph{endogenous}, not an assumption restated.

\paragraph{Own-retail destination.} The project's ground floor competes inside
the model rather than being appended outside it. Its retail square footage is
converted to POI-count-equivalent attractiveness
($A_{own} = \mathrm{sqft} / \assum{own\_retail\_sqft\_per\_equiv\_poi}$),
distributed across a fixed category mix (50\% restaurant/bar, 15\% comparison,
15\% convenience, 20\% personal services), at zero travel time from the site.
Huff's original attractiveness variable was selling area \cite{huff1963};
competitor attractiveness here is POI count because open POI data lacks floor
area, so the own-retail conversion reconciles the two units.

\paragraph{In-city shares.} For food-away and for all-retail, the in-city
capture share is the probability mass at destinations inside the city boundary,
$\theta = \sum_{j \in \mathrm{city}} \sum_m P(j,m)$. These shares are the
cross-module inputs to the meals-tax and sales-tax lines.

\paragraph{Reporting rollup.} Businesses are grouped for display by DBSCAN
(density-based spatial clustering of applications with noise) \cite{ester1996} ($\varepsilon = 150$\,m, $\mathrm{minPts} = 3$) and each
cluster is labeled by the named street intersection nearest its centroid
(from the OSM walk graph, within $\sim$450\,m), falling back to the central
member business where no named intersection is close; cluster values are sums
of member-business captures and play no role in the allocation itself.

\section{Walking activity}

\subsection{Resident walk trips and exact marginal flows}
Daily walk trips are $T = N_{res} \cdot \tau$. These all-purpose trips are
allocated over POI-weighted walk-network nodes with exponential decay and
routed along shortest paths --- deterministically, with no sampling --- giving
per-street-segment flows:
\begin{equation}
f_e \;=\; \sum_{v} T \cdot
\frac{g_v\, e^{-\beta_w t_v}}{\sum_u g_u\, e^{-\beta_w t_u}}
\cdot \mathbb{1}\!\left[ e \in \mathrm{SP}(\mathrm{site} \to v) \right],
\end{equation}
where $g_v$ is the POI weight at node $v$ and $\mathrm{SP}$ is the shortest
path. The change caused by the project \emph{is} the trips its residents
generate, so no baseline differencing (and no sampling noise) enters the
numerator.

\subsection{Baseline index and percent comparison}
The comparison baseline is a seeded, sampled, weighted betweenness index
\cite{freeman1977, brandes2001} of today's walkers: $k = 2000$
origin--destination pairs drawn with origins proportional to census-block
population within 400\,m of each node and destinations proportional to POI
weight, routed by shortest paths. Scaling the sampled counts by
$(\mathrm{total\ population\ weight} \times \tau) / k$ expresses the baseline
in the same trips units, and the trip rate $\tau$ cancels from the percentage:
\begin{equation}
\Delta_e \;=\; 100 \cdot
\frac{f_e}{n_e \cdot \bar{P}\, \tau / k}
\quad \text{(reported only where } n_e \geq 5\text{)},
\end{equation}
with $n_e$ the sampled baseline count on edge $e$ and $\bar{P}$ the total
population weight. The headline metric averages $\Delta_e$ over the ten
commercial street segments nearest the site. Without calibrated pedestrian
counts this is an index comparison, not a volume estimate; the correlation of
betweenness-type indices with observed pedestrian volumes is supported by the
network-analysis literature but is context-dependent.

\subsection{Consistency diagnostic}
\begin{equation}
\text{spending per walk trip} = \frac{\sum_j C^{walk}_j}{365\, T}.
\end{equation}
Because $T$ counts \emph{all} walking (exercise, transit access, dog walks)
while the numerator is shopping only, a plausible value is a few dollars per
average trip (implying \$5--\$15 per actual shopping trip). Values outside
\$0.25--\$30 add a consistency flag to the report.

\section{Employment ledger}

\begin{equation}
J^- = \frac{\mathrm{existing\ commercial\ sqft}}{\assum{sqft\_per\_office\_job}},
\qquad
J^+ = \frac{\mathrm{proposed\ retail\ sqft}}{\assum{sqft\_per\_retail\_job}},
\qquad
\Delta J = J^+ - J^-.
\end{equation}
Space-per-worker values (300 [200--450] office; 500 [400--700] retail, gross)
sit between dense workplace-utilization benchmarks and the building-stock
averages observable in the Commercial Buildings Energy Consumption Survey
(CBECS) \cite{cbecs}. These are site-activity screens
only; no indirect or induced employment is modeled.

\section{Fiscal module}

The module reports both of the two canonical fiscal-impact framings ---
average(per-capita) cost and marginal cost --- following Burchell \&
Listokin's classification \cite{burchell1978} and the planning literature's
standard caution that neither is universally correct for infill
\cite{kotval2006}.

\subsection{Revenue}
Real-estate (RE) tax lines use assessed value ($AV$) at the adopted rate $r$
per \$100 of value, with $p_{sqft}$ the assumed commercial value per square
foot (\assum{commercial\_value\_per\_sqft}); BPOL is Virginia's Business,
Professional and Occupational License tax on gross receipts.
\begin{align}
\text{current RE tax} &= AV_{now} \cdot r / 100 \\
AV_{proj} &= \mathrm{median}\!\left(\tfrac{AV_i}{\mathrm{units}_i}\right)
             \cdot U \;+\; \mathrm{retail\ sqft} \cdot p_{sqft}
             \qquad \text{(bounds: comp 25th--75th pct)} \\
\text{projected RE tax} &= AV_{proj} \cdot r / 100 \\
\text{RE tax increase} &= (AV_{proj} - AV_{now}) \cdot r / 100 \\
\text{meals tax} &= S_{\mathrm{food\ away}} \cdot \theta_{\mathrm{food}} \cdot r_{\mathrm{meals}} \\
\text{local sales tax} &= \left( S_{\mathrm{grocery}} + S_{\mathrm{comparison}}
                          + S_{\mathrm{convenience}} \right) \cdot \theta_{\mathrm{all}}
                          \cdot r_{\mathrm{sales}} \\
\text{personal property} &\approx N_{hh} \cdot \nu_{hh} \cdot \bar{v}_{veh}
                          \cdot r_{pp} / 100
    \qquad \text{(rough estimate)} \label{eq:ppest} \\
\text{BPOL} &\approx \mathrm{retail\ sqft} \cdot g_{sf} \cdot r_{bpol} / 100
    \qquad \text{(rough estimate)} \label{eq:bpolest}
\end{align}
Comparables are apartment-class parcels (land-use code 352) built within a
15-year lookback, pulled from the city's public assessment database with
per-unit values $AV_i / \mathrm{units}_i$; fewer than three comps degrades the
metric with a low-confidence flag. All rates ($r$, $r_{\mathrm{meals}}$,
$r_{\mathrm{sales}}$) are jurisdiction configuration pinned with source URL and
fiscal year. Virginia's 1\% local-option sales tax applies to food for home
consumption, so grocery is correctly in the taxable base; personal services are
excluded as generally non-taxable.

Personal property and BPOL revenue are \emph{rough estimates}, and are labeled
as such wherever they appear. Their tax rates are pinned city policy ---
$r_{pp} = \$4.13$ per \$100 of vehicle value and $r_{bpol} = \$0.20$ per \$100
of retail gross receipts, both from the city's adopted-budget Rates \& Levies
schedule --- but the bases are assumptions: vehicles per household
($\nu_{hh}$), average vehicle value ($\bar{v}_{veh}$), and tenant gross sales
per square foot ($g_{sf}$) have no project-specific data behind them and carry
wide bounds. When per-household personal-property budget actuals ($p_{pp}$)
are pinned, the actuals path $N_{hh} \cdot p_{pp}$ replaces
Equation~\eqref{eq:ppest}. Both estimate lines enter incremental recurring
revenue: omitting them would systematically understate revenue for taxes the
city does levy. (Car-tax relief under Virginia's Personal Property Tax Relief
Act is a fixed state block grant, so
marginal vehicles yield the city the full levy; the services BPOL
classification at \$0.27 per \$100 is slightly above the retail rate applied
here.)

\subsection{Service costs and net impact}
When the city's education transfer $E$ and school enrollment $N_{enr}$ are
pinned, costs use the \emph{school-split} model --- the per-capita-multiplier
refinement of \cite{burchell1978}: school costs follow the project's own
student estimate at the per-pupil rate, and only non-school costs are
allocated per resident. The per-pupil rate is \emph{net} of the state
education revenue $A$ the locality receives (basic aid and the education
sales tax follow average daily membership, ADM, so a new student brings that
revenue with them; only the local share is a cost to city taxpayers). In the
formulas below, $GF$ is the city's total General Fund expenditure and $P$
its citywide population.
\begin{align}
S_{K12} &= U \cdot \assum{students\_per\_unit} \\
\text{naive cost} &= N_{res} \cdot \frac{GF - E}{P}
                     \;+\; S_{K12} \cdot \frac{E - A}{N_{enr}} \\
\text{marginal cost} &= N_{res} \cdot \frac{GF - E}{P} \cdot \mu
                     \;+\; S_{K12} \cdot \frac{E - A}{N_{enr}}, \qquad
\mu = \assum{marginal\_cost\_factor} \\
\text{net}_{\,method} &= R_{inc} - \text{cost}_{\,method}
\end{align}
where $R_{inc}$ is \emph{incremental} recurring revenue (the real-estate
component is the tax \emph{increase} when the current tax is known).
The citywide identity is preserved: total allocable local cost is
$(GF - E) + (E - A) = GF - A$, i.e.\ the state-financed share of the General
Fund is excluded from local cost, exactly as a local-fiscal test requires.
If $A$ is unpinned the per-pupil rate degrades to the gross $E / N_{enr}$.
The school term appears unscaled in both framings: pupils generate tuition-contract
costs one-for-one, so a development generating few students carries
proportionally low school costs instead of the school-heavy citywide average
--- without the split, a zero-student building would still be billed average
school costs for every resident. The student generation rate is set toward the
high-rise/small-unit end of the multifamily evidence \cite{listokin2006}
because university-adjacent rental buildings house predominantly single and
roommate households rather than families. The naive framing remains an upper
bound on the non-school side (it allocates fixed citywide costs to incremental
residents); the marginal framing scales only the non-school share expected to
grow and can understate costs if capacity thresholds are crossed. The headline
range spans both framings; per-method nets are also published. If $E$ or
$N_{enr}$ is unpinned, both framings degrade to the single per-capita form
$N_{res} \cdot GF/P$ with the student count reported as a note.

\section{Assumptions}

Central values with propagated bounds; every metric lists the assumption keys
that touched it.

\begin{center}
\footnotesize
\begin{tabular}{@{}llll@{}}
\toprule
Key & Value & Range & Basis \\
\midrule
\assum{occupancy\_rate} & 0.95 & 0.90--0.97 & stabilized MF occupancy; cf.\ HVS vacancy \cite{hvs} \\
\assum{avg\_hh\_size\_multifamily} & 1.8 & 1.5--2.2 & bedroom-mix multipliers \cite{listokin2006} \\
\assum{income\_premium\_new\_construction} & 1.125 & 1.0--1.25 & screening premium, stated openly \\
\assum{ces\_scale} & 1.0 & 0.85--1.15 & CES national-average drift allowance \cite{blsces} \\
\assum{beta\_walk} (min$^{-1}$) & 0.10 & 0.05--0.15 & purpose-specific walk decay 0.09--0.11 \cite{iacono2008}; national NHTS duration decay 0.073 \cite{yang2012} \\
$\beta_d$ (min$^{-1}$, fixed) & 0.15 & --- & drive impedance decay \\
\assum{walk\_share\_neighborhood} & 0.60 & 0.40--0.80 & zero-impedance walk preference \\
\assum{walk\_share\_comparison} & 0.10 & 0.00--0.30 & comparison goods travel by car \\
\assum{walk\_share\_grocery\_entertainment} & 0.30 & 0.10--0.50 & interpolated \\
\assum{walk\_trips\_per\_resident\_day} & 1.0 & 0.5--2.0 & NHTS metro rates + walkability uplift \cite{nhts, ewing2010} \\
\assum{sqft\_per\_office\_job} & 300 & 200--450 & planning standard; cf.\ CBECS \cite{cbecs} \\
\assum{sqft\_per\_retail\_job} & 500 & 400--700 & planning standard; cf.\ CBECS \cite{cbecs} \\
\assum{own\_retail\_sqft\_per\_equiv\_poi} & 2{,}000 & 1{,}500--3{,}000 & typical inline suite size \\
\assum{commercial\_value\_per\_sqft} & \$275 & \$200--\$400 & screening; replace with comps \\
\assum{students\_per\_unit} & 0.10 & 0.05--0.15 & high-rise MF, univ.-adjacent \cite{listokin2006} \\
\assum{marginal\_cost\_factor} & 0.40 & 0.25--0.60 & share of average cost that scales \\
\assum{vehicles\_per\_household} & 1.4 & 1.0--1.8 & ACS vehicle norms, MF renters (estimate) \\
\assum{avg\_vehicle\_assessed\_value} & \$14{,}000 & \$9k--\$20k & Northern Virginia vehicle rolls (estimate) \\
\assum{retail\_sales\_per\_sqft} & \$400 & \$250--\$600 & neighborhood retail gross sales (estimate) \\
\bottomrule
\end{tabular}
\end{center}

\noindent
Structural constants: walk speed 4.8\,km/h; DBSCAN $\varepsilon=150$\,m,
$\mathrm{minPts}=3$; network extent = city boundary + 3\,km; POI study area =
boundary + 2\,mi; node population radius 400\,m; baseline sampling $k = 2000$,
fixed seed; comp lookback 15 years.

\section{Implementation decisions}

\begin{itemize}
\item \textbf{Per-business destinations.} The Huff run treats every retail POI
      as an individual destination (a single-source Dijkstra makes per-POI times
      free); DBSCAN clusters exist only as reporting rollups. Capture is therefore
      spatially resolved to business locations, and map heat layers consume real
      per-business weights rather than interpolated cluster centroids.
\item \textbf{Joint rather than nested choice.} Equation~\eqref{eq:joint} is a
      single multinomial logit over (destination, mode) pairs. A nested logit would
      relax the independence-of-irrelevant-alternatives (IIA) property across modes; the
      joint form was chosen for transparency and because screening bounds on $w_c$
      and $\beta_w$ dominate the error budget.
\item \textbf{Exact marginal flows.} Project-caused foot traffic is computed as
      the shortest-path routing of the residents' own trips, replacing an earlier
      sampled before/after betweenness difference whose noise exceeded the signal at
      this project scale. Sampling now appears only in the baseline denominator.
\item \textbf{Census-block population.} Node population weights use 2020 census
      block centroids (the finest published geography) fetched by spatial envelope,
      which also extends the baseline beyond the city line that the county-scoped
      ACS layer stopped at.
\item \textbf{Determinism.} All sampling is seeded; two evaluations on a warm
      cache produce identical metric values.
\item \textbf{Large-language-model (LLM) firewall.} Document extraction demands a verbatim
      $\leq$15-word quote per numeric field, verified in code against the document
      text; unverifiable values are nulled. Parcel identifiers are verified the
      same way.
\item \textbf{Narrative validator.} The generated report is scanned for
      numbers; each must match a metric value/bound, spec quantity, assumption, or
      deterministic note within rounding, else the draft is regenerated once and
      then hard-fails.
\item \textbf{Assessment data.} The city publishes no bulk assessment layer, so
      site values and multifamily comparables come from the public Patriot WebPro
      search interface, throttled to $\sim$1 request/s and cached; owner names are
      neither parsed nor stored.
\item \textbf{Map scales.} Heat layers are clipped to commercial-retail zoning
      polygons and colored on dollar-pinned viridis ramps (nice-number ceilings),
      so a color means the same dollars regardless of layer toggling; capture and
      walk-in capture share one scale to make their comparison meaningful.
\item \textbf{Interactive assumption adjustment.} Each metric publishes an
      exact power-law decomposition over the assumptions it depends on
      multiplicatively (or by a fixed power, e.g.\ the Engel elasticities):
      $v' = \sum_t v_t \prod_k (a_k / a_k^0)^{e_{t,k}}$. Because the joint-choice
      probabilities are independent of the demand-side assumptions, this
      client-side recompute reproduces the pipeline's own arithmetic exactly;
      the evaluation asserts $\sum_t v_t = v$ on every run. Network-model
      parameters ($\beta_w$, mode shares, own-retail attractiveness) appear in
      no term and are presented as requiring a full re-run.
\end{itemize}

\section{Known limitations and audit notes}
\label{sec:limitations}

Four limitations identified in the July 2026 audit have been addressed in this
version; their adopted resolutions and residual caveats:
\begin{enumerate}
\item \textbf{Household size (addressed).} The unit-mix-blind ACS renter
      average ($\approx 3.1$) was replaced by a multifamily bedroom-mix
      multiplier, $\bar{s} = 1.8$ [1.5--2.2] \cite{listokin2006}.
      \emph{Residual:} the multiplier assumes a typical
      studio/one-bedroom/two-bedroom mix;
      when a project's actual bedroom program is extracted it should set
      $\bar{s}$ directly.
\item \textbf{Aggregate income (addressed).} Tract income now uses the mean
      (B19025/B11001) rather than median~$\times$~households, with a median
      fallback under suppression. \emph{Residual:} tract means still smooth
      over building-level differences; the new-construction premium carries
      that uncertainty.
\item \textbf{CES income scaling (addressed).} Category spending scales by
      $r^{\varepsilon_c}$ per Equation~\eqref{eq:engel} instead of linearly.
      \emph{Residual:} the elasticities are national approximations read from
      CES quintile gradients, not locally estimated.
\item \textbf{Walk-trip rate (addressed).} $\tau$ reduced from 2.0 to 1.0
      [0.5--2.0], anchored to NHTS metro person-level walking rates with a
      walkability uplift consistent with the built-environment literature
      \cite{nhts, ewing2010}. The foot-traffic \emph{percentage} is unaffected
      ($\tau$ cancels) and walk-in dollars are independent of $\tau$; trips and
      street-flow magnitudes scale with it. \emph{Residual:} local counts
      remain the real calibration.
\item \textbf{IIA.} The joint logit imposes proportional substitution across all
      (destination, mode) pairs; a nested structure would be more standard.
\item \textbf{Transferability of $\beta_w$.} The central value (0.10) sits
      between the Twin Cities purpose-specific estimates (shopping 0.094,
      restaurant 0.093 \cite{iacono2008}) and the national NHTS duration
      decay (0.073 \cite{yang2012}). Both are fit to observed trip-length
      distributions, which conflate willingness with spatial structure
      \cite{sheppard1995}; in a competing-destinations model
      the pure impedance coefficient is plausibly gentler still. Pedestrian-
      environment quality is not separately modeled — for hostile walking
      environments, read the steep end of the band (0.15).
\item \textbf{Attractiveness proxy.} Competitor attractiveness is POI count
      (floor area being unavailable in open POI data), while own-retail
      attractiveness is derived from square footage; the conversion constant
      carries the reconciliation and its bounds.
\item \textbf{Sales-tax base share.} The taxable base applies the all-retail
      in-city share to the three taxable categories; a category-specific share
      would be marginally more precise.
\item \textbf{Foot-traffic index is uncalibrated.} Percent changes compare
      modeled flows to a sampled opportunity index, not to counted pedestrians;
      local counts would convert the layer to volumes.
\item \textbf{Fiscal method dependence.} The net range spans average-cost and
      marginal-cost framings on purpose; the midpoint of that range is a
      presentation convenience, not an estimate with independent standing.
\item \textbf{Rate-based revenue estimates.} Personal property and BPOL
      revenue use pinned city rates but assumed bases, per
      Equations~\eqref{eq:ppest} and~\eqref{eq:bpolest}; they are labeled
      rough estimates and carry the widest relative bounds on the revenue
      side. Per-household personal-property budget actuals and tenant-level
      receipts would replace them.
\end{enumerate}

\section{Data sources}

City of Fairfax GeoHub ArcGIS services (boundary, parcels, zoning); City of
Fairfax public assessment database (Patriot WebPro); City budget pages (tax
rates, General Fund, pinned with fiscal year) and the adopted-budget Rates \&
Levies schedule (personal property and BPOL rates); City of Fairfax Schools
fiscal-year 2026 budget announcement (tuition contract with Fairfax County
Public Schools and projected ADM, for the school-split cost model); Census
ACS 5-year (B01003,
B19013, B25010, B25044, B08301) and 2020 PL~94-171 block population via
TIGERweb; Census Longitudinal Employer--Household Dynamics Origin--Destination
Employment Statistics (LEHD LODES) version~8; OpenStreetMap via OSMnx
\cite{boeing2017}; Overture
Maps places; BLS Consumer Expenditure Survey 2023 \cite{blsces}; FHWA National
Household Travel Survey \cite{nhts}; Virginia Code \S 58.1-605 (local-option
sales tax).

\begin{thebibliography}{99}\small

\bibitem{huff1963} Huff, D.~L. (1963). A probabilistic analysis of shopping
center trade areas. \emph{Land Economics}, 39(1), 81--90.

\bibitem{huff1964} Huff, D.~L. (1964). Defining and estimating a trading area.
\emph{Journal of Marketing}, 28(3), 34--38.

\bibitem{wilson1971} Wilson, A.~G. (1971). A family of spatial interaction
models, and associated developments. \emph{Environment and Planning A}, 3(1),
1--32.

\bibitem{mcfadden1974} McFadden, D. (1974). Conditional logit analysis of
qualitative choice behavior. In P.~Zarembka (Ed.), \emph{Frontiers in
Econometrics} (pp.~105--142). Academic Press.

\bibitem{benakiva1985} Ben-Akiva, M., \& Lerman, S.~R. (1985). \emph{Discrete
Choice Analysis: Theory and Application to Travel Demand}. MIT Press.

\bibitem{freeman1977} Freeman, L.~C. (1977). A set of measures of centrality
based on betweenness. \emph{Sociometry}, 40(1), 35--41.

\bibitem{brandes2001} Brandes, U. (2001). A faster algorithm for betweenness
centrality. \emph{Journal of Mathematical Sociology}, 25(2), 163--177.

\bibitem{ester1996} Ester, M., Kriegel, H.-P., Sander, J., \& Xu, X. (1996). A
density-based algorithm for discovering clusters in large spatial databases
with noise. \emph{Proceedings of KDD-96}, 226--231.

\bibitem{boeing2017} Boeing, G. (2017). OSMnx: A Python package to work with
graph-theoretic OpenStreetMap street networks. \emph{Journal of Open Source
Software}, 2(12), 215.

\bibitem{burchell1978} Burchell, R.~W., \& Listokin, D. (1978). \emph{The
Fiscal Impact Handbook: Estimating Local Costs and Revenues of Land
Development}. Center for Urban Policy Research, Rutgers University.

\bibitem{kotval2006} Kotval, Z., \& Mullin, J. (2006). \emph{Fiscal Impact
Analysis: Methods, Cases, and Intellectual Debate}. Lincoln Institute of Land
Policy working paper WP06ZK2.

\bibitem{listokin2006} Listokin, D., Voicu, I., Dolphin, W., \& Camp, M.
(2006). \emph{Who Lives in New Jersey Housing? Updated New Jersey Demographic
Multipliers}. Center for Urban Policy Research, Rutgers University.

\bibitem{blsces} U.S. Bureau of Labor Statistics. \emph{Consumer Expenditure
Surveys}, 2023 tables. \url{https://www.bls.gov/cex/}

\bibitem{nhts} Federal Highway Administration. \emph{National Household Travel
Survey}, 2022. \url{https://nhts.ornl.gov/}

\bibitem{iacono2008} Iacono, M., Krizek, K., \& El-Geneidy, A. (2008).
\emph{Access to Destinations: How Close Is Close Enough? Estimating Accurate
Distance Decay Functions for Multiple Modes and Different Purposes}. Report
MN/RC 2008-11, Minnesota Department of Transportation. Table 10 estimates
negative-exponential walk decay by travel time: work 0.106, shopping 0.094,
restaurant 0.093, recreation 0.100 min$^{-1}$. Journal version: Iacono, Krizek
\& El-Geneidy (2010), \emph{Journal of Transport Geography}, 18(1), 133--140.

\bibitem{yang2012} Yang, Y., \& Diez-Roux, A.~V. (2012). Walking distance by
trip purpose and population subgroups. \emph{American Journal of Preventive
Medicine}, 43(1), 11--19. National 2009 NHTS; fitted duration decay
0.073 min$^{-1}$ ($R^2 = 0.99$); median walk trip 10 minutes; 65\% of walk
trips exceed one quarter mile.

\bibitem{sheppard1995} Sheppard, E.~S. (1995). Modeling and predicting
aggregate flows. In S. Hanson (Ed.), \emph{The Geography of Urban
Transportation} (2nd ed., pp. 100--128). Guilford Press.

\bibitem{hvs} U.S. Census Bureau. \emph{Housing Vacancies and Homeownership
(CPS/HVS)}. \url{https://www.census.gov/housing/hvs/}

\bibitem{cbecs} U.S. Energy Information Administration. \emph{Commercial
Buildings Energy Consumption Survey (CBECS)}, 2018 building characteristics
tables. \url{https://www.eia.gov/consumption/commercial/}

\bibitem{ewing2010} Ewing, R., \& Cervero, R. (2010). Travel and the built
environment: A meta-analysis. \emph{Journal of the American Planning
Association}, 76(3), 265--294.

\end{thebibliography}

\end{document}
